---
title: "Homework 10 submitted by Md Nahian Imtiaz Hasan"
author: "Stefan Konigorski"
date: "January 11, 2023"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

For all exercises, use the KiGGS dataset.

## Exercise 1: Logistic regression

Choose 1 suitable outcome variable of interest and 3 predictors, and compute a logistic regression model. Interpret the results: which predictor is associated with the outcome and what is the strength of association (odds ratio)? Also, is the model a good fit i.e. can the outcome be predicted well (look at the misclassification table for this)?
#Loading the dataset
```{r}
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
dat <- KiGGS03_06
```
# Selected variables are emaoauf.e(Emotional Problem),pw100.e(Psychic Well Being) and kw100.e(Physical Well Being)
```{r}
str(dat$sel100.e)
str(dat$pw100.e)
str(dat$emoauf.e)
table(dat$sel100.e)
table(dat$pw100.e)
table(dat$emoauf.e)


```
#Format Variable
```{r}
dat$emoauf.e <- factor(dat$emoauf.e, labels = c("Normal", "Borderline", "Noticeable"))

```
#Investigating if Physical Well being and Psychic Well Being (kw100.e and pw100.e) are associated with Emotional Problem(dat$emoauf.e), using logistic regression.
```{r}
model1 <- glm(dat$emoauf.e ~ as.numeric(dat$pw100.e) + as.numeric(kw100.e), family = binomial(link = "logit"), data = dat)
summary(model1)


```
# The idea is to set Normal to 0/no and the rest(Noticeable and Borderline)to 1/yes.After doing so,the model will be recomputed.  
```{r}
dat$emoauf.e <- factor(dat$emoauf.e, labels = c("no", "yes", "yes"))
table(dat$emoauf.e)
dat$emoauf.e <- droplevels(dat$emoauf.e)

# Compute the odds ratio:
exp(coef(model1))
library(jtools)
jtools::summ(model1, exp = T, confint = T, model.fit = F, digits = 3)
```
#Recomputing the model
```{r}
model1997 <- glm(dat$emoauf.e ~ as.numeric(dat$pw100.e) + as.numeric(kw100.e), family = binomial(link = "logit"), data = dat)
summary(model1997)
```
```{r}
# Compute the odds ratio:
exp(coef(model1997))
# Compute 95% confidence intervals of the odds ratio
# install.packages("jtools")
library(jtools)
jtools::summ(model1, exp = T, confint = T, model.fit = F, digits = 3)
```
## Exercise 2: Poisson regression

Predict the amount of measles vaccinations (Maanzahl) by the number of siblings (e006B1), the sex and age of the children (sex, age2), place of residence (STALA, OW) and the monthly household income (e093), using a Poisson regression model. Interpret the results. Which variables are associated with the outcome? Is the model a good fit to the data?

```{r}
str(dat$Maanzahl)
str(dat$e006B1)
str(dat$sex)
str(dat$age2)
str(dat$STALA)
str(dat$OW)
str(dat$e093)
table(dat$Maanzahl)
table(dat$e006B1)
table(dat$sex)
table(dat$age2)
table(dat$STALA)
table(dat$OW)
table(dat$e093)
```
```{r}
dat$Maanzahl<- as.numeric(as.character(dat$Maanzahl))
dat$e006B1<- factor(dat$e006B1, labels = c("no sibling", "1 sibling", "2 or more siblings"))
dat$sex   <- factor(dat$sex, labels = c("boys", "girls"))
dat$age2  <- factor(dat$age2, labels = c("0-1y", "2-3y", "4-5y", "6-7y", "8-9y", "10-11y", "12-13y", "14-15y", "16-17y"))
dat$OW    <- factor(dat$OW, labels = c("east", "west"))
dat$STALA <- factor(dat$STALA, labels = c("rural", "small_city", "medium_city", "urban"))
```
```{r}
# Look at variables
table(dat$E085z01)
table(dat$e006B1)
table(dat$sex)
table(dat$age2)
table(dat$STALA)
table(dat$OW)
table(dat$e093)

```
#Computing Mean And Variance
```{r}
hist(dat$Maanzahl)
mean(dat$Maanzahl, na.rm = TRUE)
var(dat$Maanzahl, na.rm = TRUE)

#Mean and Var are not that far apart.
```
```{r}
# Poisson Regression: first predict dat$Maanzahl only by sex and age:
fit2000 <- glm(dat$Maanzahl ~ dat$sex+dat$age2, family = poisson(link = "log"), data = dat)
# results:
summary(fit2000)
exp(coef(fit2000))

# Direcly using the summ function:
library(jtools)
jtools::summ(fit2000, exp = T, confint = T, model.fit = F, digits = 3)
```
```{r}
# Predicted frequencies by the model:
pred_2000 <- predict(fit2000, type="response")

# Compare these with the observed frequencies:
predictions_2000 <- data.frame(obs = dat$Maanzahl[1:17640 %in% names(pred_2000)], pred_2000 = pred_2000)
cor(predictions_2000$obs, predictions_2000$pred_2000)
plot(predictions_2000$obs, predictions_2000$pred_2000)
```
```{r}
#first predict dat$Maanzahl only STALA and OW:
fit2001 <- glm(dat$Maanzahl ~ dat$STALA+dat$OW, family = poisson(link = "log"), data = dat)
# results:
summary(fit2001)
exp(coef(fit2001))

# Direcly using the summ function:
library(jtools)
jtools::summ(fit2001, exp = T, confint = T, model.fit = F, digits = 3)
```
```{r}
# Poisson Regression: first predict dat$Maanzahl only by the number of siblings:
fit2002 <- glm(dat$Maanzahl ~ dat$e006B1, family = poisson(link = "log"), data = dat)
# results:
summary(fit2002)
exp(coef(fit2002))

# Direcly using the summ function:
library(jtools)
jtools::summ(fit2002, exp = T, confint = T, model.fit = F, digits = 3)
```

```{r}
#Poisson Regression: first predict dat$Maanzahl only by the the monthly household income (e093)
fit2003 <- glm(dat$Maanzahl ~ dat$e093, family = poisson(link = "log"), data = dat)
# results:
summary(fit2003)
exp(coef(fit2003))

# Direcly using the summ function:
library(jtools)
jtools::summ(fit2003, exp = T, confint = T, model.fit = F, digits = 3)
 
```
```{r}
#COMPUTING ALL OF THEM TOGETHER
fit2010 <- glm(dat$Maanzahl ~ dat$e093+dat$e006B1+dat$STALA+dat$OW+dat$sex+dat$age2, family = poisson(link = "log"), data = dat)
# results:
summary(fit2010)
exp(coef(fit2010))

# Direcly using the summ function:
library(jtools)
jtools::summ(fit2010, exp = T, confint = T, model.fit = F, digits = 3)

```
## Exercise 3: Negative Binomial regression (optional)

Predict the amount of measles vaccinations (Maanzahl) by the number of siblings (e006B1), the sex and age of the children (sex, age2), place of residence (STALA, OW) and the monthly household income (e093), using a Negative Binomial regression model. Interpret the results. Which variables are associated with the outcome? Is the model a good fit to the data?
