---
title: "Homework 12 submitted by Md Nahian Imtiaz Hasan"
author: "Stefan Konigorski"
date: "January 25, 2023"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Linear mixed model for hierarchical data

Load the HRS dataset and do the following preparations:

```{r}
library(readr)
dat_link <- url("https://www.dropbox.com/s/z281ijk81ohyg9v/hrs.RData?dl=1")
load(dat_link)



# Preparations
# rename variable at first timepoint
names(hrs)[5] <- "R1IMRC"
# define variable at second timepoint - Integrate the variables R2AIMRC10 and R2HIMRC20, which both capture different subgroups of the sample at the second timepoint.
table(is.na(hrs$R2AIMRC10)) # -> only ~7000 of ~36000 observations
table(is.na(hrs$R2HIMRC20)) # -> only ~10000 of ~36000 observations
table(!is.na(hrs$R2HIMRC20) & !is.na(hrs$R2AIMRC10)) # -> no person has a value at both variables, i.e. the variables complement each other
hrs$R2IMRC <- rowSums( cbind (hrs$R2AIMRC10, hrs$R2HIMRC20), na.rm=TRUE)
hrs$R2IMRC[!(!is.na(hrs$R2HIMRC20) | !is.na(hrs$R2AIMRC10))] <- NA
table(is.na(hrs$R2IMRC))
# Rescale R1IMRC and R2IMRC onto the same scale as all other variables (measures 3-12 were on a scale of 1-10, measures 1-2 on the scale 1-20)
hrs$R1IMRC <- hrs$R1IMRC/2
hrs$R2IMRC <- hrs$R2IMRC/2

# Extract variables for first investigation
data_IMRC <- data.frame(HHID = hrs$HHID, PN = hrs$PN, HHID_PN = paste(hrs$HHID, hrs$PN, sep = "_"),
                        T1 = hrs$R1IMRC, T2 = hrs$R2IMRC, T3 = hrs$R3IMRC, T4 = hrs$R4IMRC, T5 = hrs$R5IMRC, T6 = hrs$R6IMRC,
                        T7 = hrs$R7IMRC, T8 = hrs$R8IMRC, T9 = hrs$R9IMRC, T10 = hrs$R10IMRC, T11 = hrs$R11IMRC, T12 = hrs$R12IMRC)
summary(data_IMRC$T1)
```

Answer the following questions: 

a) How many observations are there in the dataset?
#Answer: There are 36167 observations in the dataset. 

b) How many households are there in the dataset (observations are clustered in households)?
#Answer: There are 36167 households in the dataset. 
c) Now, compute a linear regression model to predict T2 by T1, i.e. the cognitive performance at the second timepoint by the cognitive performance at the first timepoint.
```{r}
tapply(data_IMRC$T2, data_IMRC$T2, mean, na.rm = TRUE)
modela <- lm(T2 ~ T1, data = data_IMRC )
summary(modela)
predictiona <- predict(modela, newdata = data_IMRC)
summary(predictiona)
head(predictiona)
# beta= 0.53446
#the same value is predicted at T1 for all persons, at T2 the value from T1 + 0.53446 is predicted for all persons.

# Model fit
AIC(modela)
plot(data_IMRC$T1, data_IMRC$T2)
abline(modela)
```
Compute two linear mixed models, one with random intercept, one with random intercept+slope to predict T2 by T1. Compare the results, i.e. the estimated effect of T1 on T2. Are they similar, or are they very different?
#One with HHID_PN as the Random intercept
```{r}

library(nlme)

modelb <- lme(T2 ~ T1, random = ~1|HHID_PN, na.action = na.exclude, data = data_IMRC)
summary(modelb)
pred.modelb <- predict(modelb)
# Look at the predicted values:
pred.modelb[1:20]
# Visualization of the predicted values
ggplot(data_IMRC, aes(x = T1, y = T2)) + 
  stat_summary(aes(y = pred.modelb), fun.y = mean, colour = "green", geom = "line")
# -> Obviously,a different curve as in linear regression

```
#Compute another linear mixed model-one with random intercept+slope to predict T2 by T1.
```{r}

modelc <- lme(T2 ~ T1, random = ~PN|HHID_PN, na.action = na.exclude, data = data_IMRC)
summary(modelc)
pred.modelc <- predict(modelc)
# Look at the predicted values:
pred.modelc[1:20]
# Visualization of the predicted values
ggplot(data_IMRC, aes(x = T1, y = T2)) + 
  stat_summary(aes(y = pred.modelc), fun.y = mean, colour = "green", geom = "line")
# -> I have run into an error in the final question.

```
Compare the results, i.e. the estimated effect of T1 on T2. Are they similar, or are they very different?