---
title: "Homework 13 submitted by Md Nahian Imtiaz Hasan"
author: "Stefan Konigorski"
date: "February 1, 2023"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Survival analysis

Choose one other provided dataset in the survival package (or any other suitable dataset) which contains time-to-event data. Use the survfit function to compute the Kaplan-Meier estimate of the survival function and plot it. Then choose a variable with 2 or more categories in the dataset and perform a log-rank test. Interpret the results: is there are difference in survival between the groups?

```{r}
# Loading the survival package and the lung cancer data
library(survival)
lung
data(cancer, package="survival")
#Formatting the data for analysis
lung$sex <- factor(lung$sex, labels = c("male", "female"))
lung$status <- lung$status - 1


# Computing the Kaplan-Meier estimate of the survival function
fit001 <- survfit(Surv(time, status) ~ 1, data = lung)

# Plotting the Kaplan-Meier estimate
plot(fit001, xlab = "Time (days)", ylab = "Survival Probability")

# Performing a log-rank test
fit002 <- survfit(Surv(time, status) ~ sex, data = lung)
plot(fit002, xlab = "Time (days)", ylab = "Survival Probability")
summary(fit002, test = "logrank")


# Computing the Kaplan-Meier estimate of the survival function for weight loss
fit003 <- survfit(Surv(wt.loss, status) ~ 1, data = lung)
plot(fit003,xlab = "Weight Loss",ylab = "Survival Probability")
#Not a good idea TO use weight loss variable in my opinion.
```
```{r}
#Computing and plotting the Kaplan-Meier estimate of the survival function

fit004 <- survfit(Surv(meal.cal, status) ~ 1, data = lung)
plot(fit004,xlab = "Calories Consumed",ylab = "Survival Probability")
# Performing a log-rank test
fit005 <- survfit(Surv(meal.cal, status) ~ sex, data = lung)
plot(fit005, xlab = "Calories Taken", ylab = "Survival Probability")
summary(fit005, test = "logrank")
fit005
```

```{r}

```
