---
title: "Homework 5 submitted by Md Nahian Imtiaz Hasan "
author: "Stefan Konigorski"
date: "November 16, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Probability distributions

Explore the functions *rnorm*, *rt*, *runif*, *rbinom* in R that allow you to generate random numbers from the normal, t-, uniform, and binomial distribution. Compute them with different values, and inspect histograms to visualize their distribution.
```{r}
pnorm(21.3, mean=60, sd=28)

```
#rnorm histogram
```{r}
x <- rnorm(5000, mean=120, sd=25)
hist(x, probability=TRUE)

```
#rt function
```{r}
y<-rt(2000,5,1)
hist(y,probability = TRUE)
```
#runif function
```{r}
z<-runif(96, min=25, max=100)
hist(z,probability = TRUE)
```
## Exercise 2: Odds ratio

In the KiGGS dataset:

#a)  Compute the proportion of mothers that had hypertension during pregnancy. Use the variable 'e0155' which has values "Ja" (yes), "Nein" (No) and "Weiß nicht" (don't know).

```{r}
hype<-KiGGS03_06$e0155
summary(hype)
plot(hype)
```

```{r}
summary(hype)
202/17640
```
#b) Create a new variable that is 0 or 1 depending on whether the children are small or tall (think of a good way how to do this) based on the variable 'groeB' (body height). 
```{r}
height<-KiGGS03_06$groeB
summary(height)
```

c) Then compute the odds ratio that the mother had hypertension during pregnancy (e0155 == "Ja" (yes), versus e0155 == "Nein" (no)) of tall vs. small children.

## Exercise 3 (optional): Confidence intervals

Look at the hypertension variable from exercise 2. Use the binom::binom.confint and the questionr::odds.ratio functions to compute the estimates of the proportion and odds ratio as well as their confidence intervals. (you need to download and load these packages at first).

## Exercise 4 (optional, advanced): Bootstrap

Adapt the bootstrap implementation in R_5b_estimation_bootstrap.Rmd to compute the bootstrap estimate of the standard error of the variance of a normally-distributed and a t-distributed variable. Are they similar?