---
title: "Homework 7 submitted by Md Nahian Imtiaz Hasan"
author: "Stefan Konigorski"
date: "November 30, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Preparations

Load KiGGS data:

```{r}
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
kiggs <- KiGGS03_06
```

## Exercise 1: Multiple imputation

a) In the KiGGS dataset, choose 2 metric variables of your choice.
```{r}
x<- kiggs$e024z1 #Cold Infection
y<-kiggs$E014z #Weight gain in childbirth
head(y)
```
b) Compute the correlation between these two variables

```{r}
summary(x)
summary(y)
table(!is.na(as.numeric(as.character(x))))
table(!is.na(as.numeric(as.character(y))))

table(!is.na(as.numeric(as.character(x))) & !is.na(as.numeric(as.character(y))))
hist(as.numeric(as.character(x)))
hist(as.numeric(as.character(y)))
```
c) Perform a multiple imputation using the mice function with the 'pmm' method and otherwise default settings.
d) Compute the correlation again on the multiple imputed datasets.
