---
title: "Homework 13 - tutorial"
author: "Thomas Gaertner"
date: "February 7, 2023"
output:
  html_document: 
    toc: true
---

## Exercise 1: Survival analysis

### Task

> Choose one other provided dataset in the survival package (or any other suitable dataset of interest) which contains time-to-event dataset. Use the survfit function to compute the Kaplan-Meier estimate of the survival function and plot it. Then choose a variable with 2 or more categories in the dataset and perform a log-rank test. Interpret the results: is there are difference in survival between the groups?

```{r include=FALSE, show=FALSE}
library(survival)
library(survminer)
library(dplyr)
```

### Example Solution 1: NCCTG Lung Cancer Data
Survival in patients with advanced lung cancer from the North Central Cancer Treatment Group. Performance scores rate how well the patient can perform usual daily activities.
We are interested in `status` after `time` for different `sex`.

#### Data
```{r}
# preprocessing
lung$sex <- as.factor(lung$sex)
levels(lung$sex) <- c("male","female")
str(lung)
```

#### Kaplan-Meier Plots
```{r}
KM_lung <- survfit(Surv(time, status) ~ sex, data = lung)

plot(KM_lung, main = expression(paste("Kaplan-Meier-estimate ", hat(S)(t), " stratified by sex")), ylab = "Survival function", xlab = "Time", lwd=2, col=1:2)
legend(x="topright", col = 1:2, lwd = 2, legend=c("male", "female"))


ggsurvplot(KM_lung, data = lung, pval = TRUE)
```


#### Log-Rank Test

To compare the interventions, we are using the function `survdiff` to perform a log-rank test.
```{r}
# perform log-rank test:
survdiff(Surv(time, status) ~ sex, data = lung)
# Interpretation: As the p-value is smaller then the a significant level of 0.01, we can reject the null hypothesis and assume, that the time to death differs between the male and female.
```

#### Cox -Proportional-Hazard Model

```{r}
cox <- coxph(Surv(time, status) ~ sex, data = lung)
summary(cox)
```

*Interpretation*

1) *Significant*: The summary gives several p-values for overall significant in the end. As all p-values are less than the significant level of 0.01, we can reject the null hypothesis and assume that the model is significant.
2) *Significant of Coefficients*: The predictor has a p-value less than 0.01, so sex as a significant influence on the outcome. 
3) *Coefficients*: The coefficient for `sexfemale` is negative, so the risk of death is smaller compared to male.
4) *Hazard Ratios (`exp(coef)`)*: The hazard ratios gives the size of the difference. In this case being female reduces the risk of death of 41%.


### Example Solution 2: Breast cancer data set
The `rotterdam` data set includes 2982 primary breast cancers patients whose data whose records were included in the Rotterdam tumor bank.

In this data set, we analyse the survival time `dtime` for the outcome `death` for tumor size, a factor with levels (<=20, 20-50, >50). We will look at a significant level of $\alpha \leq 0.01$.

#### Data
```{r}
str(rotterdam)
#help(rotterdam)
```

#### Kaplan-Meier Plots

```{r}
KM_rotterdam <- survfit(Surv(dtime, death) ~ size, data = rotterdam)
ggsurvplot(KM_rotterdam, data = rotterdam, pval = TRUE)
```

#### Log-Rank Test

```{r}
# perform log-rank test:
survdiff(Surv(dtime, death) ~ size, data = rotterdam)
# Interpretation: As the p-value is smaller then the a significant level of 0.01, we can reject the null hypothesis and assume, that the time to death differs between the tumor sizes.
```

#### Cox-Proportional-Hazard Model

```{r}
cox <- coxph(Surv(dtime, death) ~ size, data = rotterdam)
summary(cox)
```
*Interpretation*
1) *Significant*: The summary gives several p-values for overall significant in the end. As all p-values are less than the significant level of 0.01, we can reject the null hypothesis and assume that the model is significant.
2) *Significant of Coefficients*: Each predictor has a p-value. Both are less than 0.01, so they are significant. 
3) *Coefficients*: Both coefficients for `size20-50` and `size>50` are positive, so the risk of death is greater compared to `size<20`.
4) *Hazard Ratios (`exp(coef)`)*: The hazard ratios gives the size of the difference. For example, `size20-50` has a 95% higher risk of death.

### Example Solution 3: diabetic data set

Do a survival analysis of the 'diabetic' dataset in the survival package, which contains data on the outcome time-to-visual loss, in a sample of patients with diabetic retinopathy that were treated with two types of laser treatment.


#### Data

```{r}
dat <- diabetic

# look at data:
#?diabetic
str(dat)
```

#### Kaplan-Meier Plots


```{r}
KM1 <- survfit(Surv(time, status) ~ 1, data = dat)
# Look at the plot:
plot(KM1, main = expression(paste("Kaplan-Meier-estimate ", hat(S)(t), " with 95% CI")), ylab = "Survival function", xlab = "Time")
```

In the next step, we could investigate the influence of the intervention. For that, we are creating our survival analysis stratified by the intervention `laser`. 

```{r}
# Stratify this by treatment:
KM2 <- survfit(Surv(time, status) ~ laser, data = dat)
plot(KM2, main = expression(paste("Kaplan-Meier-estimate ", hat(S)(t), " stratified by laser")), ylab = "Survival function", xlab = "Time", lwd=2, col=1:2)
legend(x="topright", col = 1:2, lwd = 2, legend=c("xenon", "argon"))
```

```{r}
ggsurvplot(KM2, data = dat, pval = TRUE)
```


#### Log-Rank Test

To compare the interventions, we are using the function `survdiff` to perform a log-rank test.
```{r}
# perform log-rank test:
survdiff(Surv(time, status) ~ laser, data = dat)
# Interpretation: no effect of treatment on the outcome = time-to-visual loss.
```

#### What is the probability after 18 Month?
Possible Question could be, what is the probability of visual loss after 18 months?
```{r}
summary(KM2, times = 18)
# Interpretation: The probability for survival is around 76% for xenon and 77% for argon after 18 month. 
```

#### Cox-Proportional-Hazard Model
```{r}
cox <- coxph(Surv(time, status) ~ laser, data = dat)
summary(cox)
```
> Interpretation:
> The log rank test had a p-value of 0.9, so we cannot reject the null-hypthesis and we have to assume that the survival curve does not depend on the laser.
> If we apply a cox-proportinal hazard model, we can see, that the coeffizient itself is slightly positive, so a risk for visual loss is slightly highy using argon laser, BUT this is not significant as the p-value is 0.89. 


### Example Solution 4: Ovarian data set

In this data set, we have RCT data from a ovarian cancer trial. We have the time-to-event variable(`futime`), the censor status (`fustst`), the treatment group (`rx`) and other covariates.  

#### Data
```{r}
str(ovarian)
#help(ovarian)
```
#### Kaplan-Meier Plots

```{r}
KM_ovarian <- survfit(Surv(futime, fustat) ~ 1, data = ovarian)

# Look at the plot
ggsurvplot(KM_ovarian, data = ovarian, pval = TRUE)
```


```{r}
# Stratify this by treatment:
KM_ovarian_2 <- survfit(Surv(futime, fustat) ~ rx, data = ovarian)

ggsurvplot(KM_ovarian_2, data = ovarian, pval = TRUE)
```

#### Log-Rank Test
```{r}
# perform log-rank test:
survdiff(Surv(futime, fustat) ~ rx, data = ovarian)
# Interpretation: no effect of treatment on the outcome = time-to-event loss.
```


